### 2024-03-10 manuscript.Rmd
Start working on the project

---
title             : "Formal Model of the Set-Point Theory of Subjective Well-Being"
shorttitle        : "Set-Point Theory of Subjective Well-Being"

author: 
  - name          : "Alp Kaan Aksu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Leopoldstraße 13"
    email         : "a.aksu@campus.lmu.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "LMU Munich"

authornote: |
  Matriculation Number: 12607598

abstract: |
  There is research that shows...

  Here, we model the set point theory of subjective well being

keywords          : ["happiness", "subjective well-being", "hedonic treadmill", "set-point theory"]
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

source("VAST/VAST.R")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r construct-definitions, include=FALSE}
swb <- construct("SWB")
swb_set_point <- construct("SWBSetPoint")
adaptation_rate <- construct("AdaptationRate")
event <- construct("Event")
```

# Deneme
```{r vast-display}
VAST(swb, swb_set_point, adaptation_rate, event) |>
  rel(swb_set_point %-c->% swb) |>
  rel(adaptation_rate %-c->% swb) |>
  rel(event %-c->% swb) |>
  analysts("Alp") |> fimm() |>
  render_vast()
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::