### 2024-03-16 material/variable_definitions.Rmd

| **Construct** | **Scale Level** | **Range** | **Anchors** |
| --------- | ----------- | ----- | ------- |
| Subjective Well-Being | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
| Subjective Well-Being Set Point | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
| Appraisal of Life Event | continuous | [-1; 1] | -1 = negative \newline 0 = neutral \newline 1 = positive |