### 2024-03-17 manuscript.Rmd
Continue working on the theory

---
title             : "Formal Model of the Set-Point Theory of Subjective Well-Being"
shorttitle        : "Set-Point Theory of Subjective Well-Being"

author: 
  - name          : "Alp Kaan Aksu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Leopoldstraße 13"
    email         : "a.aksu@campus.lmu.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "LMU Munich"

authornote: |
  Matriculation Number: 12607598

abstract: |
  There is research that shows...

  Here, we model the set point theory of subjective well being

keywords          : ["happiness", "subjective well-being", "hedonic treadmill", "set-point theory"]
wordcount         : "X"

bibliography      : ["r-references.bib", "references.bib"]

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

source("VAST/VAST.R")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

One of the most important questions we might attempt to answer is how to live our lives and what makes life good. Trying to maximize our happiness is the naive answer we easily arrive at, we want to feel good, but *happiness* is an ambiguous word. Although it seems like all people tend to evaluate their lives and think about *happiness* [@veenhoven11HowUniversal2010], the understanding of *happiness*, its use and connotations vary across time and place [@oishiConceptsHappinessTime2013]. @dienerSubjectiveWellbeing1984 suggests the use of a scientific term: *subjective well-being* (SWB).

## Subjective Well-Being
According to @dienerSubjectiveWellbeing1984, SWB is the subjective global evaluation of life and consists of three components: (1) life satisfaction, (2) positive affect, (3) negative affect.

## The Phenomenon: Hedonic Treadmill

The observation, that people go back to a stable level of SWB after (major) life events is not new. @helsonAdaptationlevelTheoryExperimental1964 formulated the more general Adaptation Level Theory, stating that how people perceive stimuli is influenced by their past experiences. @brickman1971adaptation applied this idea to SWB for the first time and coined the term *hedonic treadmill*, which is still commonly used to refer to the phenomenon of returning to a base SWB level after life events [e.g. @czaikaMigrationAspirationsAre2014, manciniSteppingHedonicTreadmill2011], and @brickmanLotteryWinnersAccident provided first evidence for it.

## Our Goal

We aim to formalize the Set-Point Theory of SWB as described by @dienerHedonicTreadmillRevising2006a, which explains the phenomenon of hedonic treadmill and tries to account for individual differences.

# Robustness of the phenomenon

@luhmannSubjectiveWellbeingAdaptation2012 has done a meta-analysis on adaptation to life events focusing on SWB. Evaluating the robustness of the studies focusing on SWB and adaptation is difficult, since it is impossible to manipulate life events experimentally and cross-sectional studies have a lack of internal validity [@luhmannSubjectiveWellbeingAdaptation2012]. All of the following statements are based on the meta-analysis by @luhmannSubjectiveWellbeingAdaptation2012.

## Empirical Evidence

Studies focusing on the hedonic treadmill can not be adequatly characterized by single effect sizes, since the temporal development of SWB is of interest. So @luhmannSubjectiveWellbeingAdaptation2012 used a method that takes time into account. For each life event, they fitted a logarithmic model using the effects sizes reported in original studies. An adaptation is characterized by a change parameter, which is in the opposite direction of the initial reaction. For example, having a positive reaction (higher SWB) after a life event, then a negative change in SWB, indicates that an adaptation is taking place.

Adaptations were observed for all life events, which means that the hedonic treadmill can be considered a robust phenomenon, but reduced internal validity due to the design of these studies makes it hard to evaluate.

## Generalizability

We evaluate the generalizability of the findings guided by the UTOS framework  [@cronbach1982designing]. A strict application of the UTOS framework to the phenomenon at hand is not possible because of the non-experimental nature of the studies.

### Units
In a lot of studies, the majority of the participants were females (65,4% across all events; for some life events like bereavement and divorce even 70+). More than half of the studies were from the US.

### Treatment
Studies focus on different major life events, 8 of them were included in the meta-analysis.

### Outcome
Using affective (AWB) and cognitive well being (CWB) measure (for some life events), studies used different measures for different variables, e.g. for the measurement of CWB in the context of marriage, relationship satisfaction was used.

### Setting
aedasd

Participants from diverse backgrounds are necessary for generalizability. Mixed (positive or negative) set of life event were analyzed, which is a good first sign for generalizability to different life events.

# The Prototheory

## Constructs

*TODO*

- SWB: Definiton by Diener (include subtypes in the theoretical model but not in)
  - AWB:
  - CWB:
- SWBSetPoint: SWB people return to
- AdaptationRate: How fast the adaptation happens
  - AdaptationRatePerson: Personenvariable
  - AdaptationRateEvent: Hängt von InitialReaction ab
- InitialReaction: Change in SWB directly after / at the time of the LifeEvent
- LifeEvent: Status change

# Model

The central idea of

# Implementation

We made small changes to the theoretical model, in order to make it suitable for an implementation. We are making the following assumptions:

- The constructs of *Life Event* and *Initial Reaction* can be reduced to *Initial Reaction*, since only the latter is a part of the (following) relevant psychological processes.
-

*VAST display of the model here*

```{r child="./material/variable_definitions.Rmd"}
```

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::