### 2024-03-17 material/statements.Rmd

| Identifier | Citation                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               | Reference                                                            | Relationship?                      | Comments?                                                                      |     |
| ---------- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ | -------------------------------------------------------------------- | ---------------------------------- | ------------------------------------------------------------------------------ | --- |
| A          | A notable finding was that the initial reaction and the rate of adaptation varied considerably between different life events.                                                                                                                                                                                                                                                                                                                                                                          | Luhmann et al. (2012)                                                | Appraisal  - c → Adaptation Rate   | Valence of the life event changes the adaptation rate.                         |     |
| A1         | For instance, the initial reaction to marriage was positive, but subsequent adaptation was fast and completed after 2 years on average (Lucas et al., 2003).                                                                                                                                                                                                                                                                                                                                           |                                                                      |                                    | positive valence: fast adaptation                                              |     |
| A2         | In contrast, the rate of adaptation was much slower for negative events such as the onset of disability (Lucas, 2007b), widowhood (Lucas et al., 2003; Specht, Egloff, & Schmukle, 2011), and divorce (Lucas, 2005).                                                                                                                                                                                                                                                                                   |                                                                      |                                    | negative valence: slow adaptation                                              |     |
| A3         | The findings were perhaps most striking for unemployment: After the beginning of unemployment, mean levels of SWB were significantly below baseline, even if people became reemployed (Lucas et al., 2004). Repeated unemployment spells aggravate this effect (Luhmann & Eid, 2009).                                                                                                                                                                                                                  |                                                                      |                                    |                                                                                |     |
| B          | Together, these studies indicate that major life events can have strong effects on SWB and that the strength of these effects varies, depending on the life events considered.                                                                                                                                                                                                                                                                                                                         |                                                                      |                                    |                                                                                |     |
| B1         | major life events can have strong effects on SWB                                                                                                                                                                                                                                                                                                                                                                                                                                                       |                                                                      | Life Event -c→ SWB                 |                                                                                |     |
| B2         | the strength of these effects varies, depending on the life events considered                                                                                                                                                                                                                                                                                                                                                                                                                          |                                                                      | Life Event -c→ intensity of change |                                                                                |     |
| C          | SWB relates to how people feel and think about their lives.                                                                                                                                                                                                                                                                                                                                                                                                                                            | Indirectly from Diener (1984) via Luhmann et al. (2012)              | n                                  |                                                                                |     |
| D          | AWB and CWB are distinct constructs                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | Indirectly from Lucas, Diener & Suh (1996) vie Luhmann et al. (2012) |                                    |                                                                                |     |
| D1         | Affective well-being (AWB) refers to the presence of pleasant affect (e.g., feelings of happiness) and the absence of unpleasant affect (e.g., depressed mood).                                                                                                                                                                                                                                                                                                                                        |                                                                      | n                                  |                                                                                |     |
| D2         | Cognitive well-being (CWB) refers to the cognitive evaluation of life overall (i.e., global life satisfaction) as well as of specific life domains (e.g., job satisfaction or marital satisfaction).                                                                                                                                                                                                                                                                                                   |                                                                      | n                                  |                                                                                |     |
| E          | It is therefore plausible that other external life circumstances, such as life events, have differential effects on AWB and CWB and that adaptation of AWB and CWB does not occur at the same rate.                                                                                                                                                                                                                                                                                                    | Luhmann et al. (2012)                                                |                                    | Having AWB and CWB separated in the model could be a good idea                 |     |
| F          | Life events have been examined from two major perspectives (Filipp & Aymanns, 2009): a stress perspective and a developmental perspective.                                                                                                                                                                                                                                                                                                                                                             | Luhmann et al. (2012)                                                |                                    |                                                                                |     |
| F1         | From the stress perspective, life events are viewed as specific types of stressors (e.g., Park, 2010; Segerstrom & Miller, 2004). Stressful life events are all events that significantly disturb the daily routine (Turner & Wheaton, 1997). This definition explicitly includes desirable events such as marriage or vacation (Holmes & Rahe, 1967).                                                                                                                                                 |                                                                      | Life Event -c→ Stress              | The stress perspective is easier to model.                                     |     |
| F2         | By contrast, minor stressors such as daily hassles and uplifts (Kanner, Coyne, Schaefer, & Lazarus, 1981) are distinct from stressful life events. From the developmental perspective, life events are viewed as specific transitions. Transitions are defined as a “discontinuity in a person’s life space of which he is aware and which requires new behavioural responses” (Hopson & Adams, 1976, p. 24).                                                                                          |                                                                      |                                    |                                                                                |     |
| F3         | For the purpose of this metaanalysis, we developed a narrower definition of life events that integrates elements from both perspectives. According to our working definition, life events are time-discrete transitions that mark the beginning or the end of a specific status. A status is a nominal variable with at least two levels.                                                                                                                                                              | Luhmann et al. (2012)                                                | Life Event -c→ Status              |                                                                                |     |
| G          | the process perspective explicitly predicts the trajectory of SWB over time within individuals                                                                                                                                                                                                                                                                                                                                                                                                         | Luhmann et al. (2012)                                                |                                    |                                                                                |     |
| G1         | The adaptation process is initiated by an external stimulus (e.g., a major life event) that causes a physiological or psychological response (e.g., decreased SWB). Over time, the responsiveness diminishes and the level of SWB returns to its preevent level.                                                                                                                                                                                                                                       | Luhmann et al. (2012)                                                |                                    | IMPORTANT<br><br>This description can be used to model the adaptation process! |     |
| H          | Most major life events are—at least to some extent— controllable and predictable and can therefore be anticipated. This anticipation might cause a specific hedonic reaction even before the event occurred: If the spouse is terminally ill, the hedonic reaction to bereavement starts long before the spouse actually dies. These so-called anticipatory or lead effects can be observed months or even several years before the occurrence of the event (Clark, Diener, Georgellis, & Lucas, 2008) | Luhmann et al. (2012)                                                |                                    |                                                                                |     |