### 2024-03-19 material/statements.Rmd

- ”MA” (meta-analysis): Luhmann et al. (2012)
- “TP” (theory paper): Diener et al. (2006)
- “OI” (original idea): Ideas / assumptions I added myself.
- “OI(MA/TP/MA+TP)”: Original ideas I extracted from things mentioned  in the meta-analysis or theory paper. It is also possible to combine insights from both papers.

| ID  | Citation                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               | Reference                                    |
| --- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ | -------------------------------------------- |
| A   | A notable finding was that the initial reaction and the rate of adaptation varied considerably between different life events.                                                                                                                                                                                                                                                                                                                                                                          | MA                                           |
| A1  | For instance, the initial reaction to marriage was positive, but subsequent adaptation was fast and completed after 2 years on average (Lucas et al., 2003).                                                                                                                                                                                                                                                                                                                                           | MA                                           |
| A2  | In contrast, the rate of adaptation was much slower for negative events such as the onset of disability (Lucas, 2007b), widowhood (Lucas et al., 2003; Specht, Egloff, & Schmukle, 2011), and divorce (Lucas, 2005).                                                                                                                                                                                                                                                                                   | MA                                           |
| A3  | The findings were perhaps most striking for unemployment: After the beginning of unemployment, mean levels of SWB were significantly below baseline, even if people became reemployed (Lucas et al., 2004). Repeated unemployment spells aggravate this effect (Luhmann & Eid, 2009).                                                                                                                                                                                                                  | MA                                           |
| B   | Together, these studies indicate that major life events can have strong effects on SWB and that the strength of these effects varies, depending on the life events considered.                                                                                                                                                                                                                                                                                                                         | MA                                           |
| B1  | major life events can have strong effects on SWB                                                                                                                                                                                                                                                                                                                                                                                                                                                       | MA                                           |
| B2  | the strength of these effects varies, depending on the life events considered                                                                                                                                                                                                                                                                                                                                                                                                                          | MA                                           |
| C   | SWB relates to how people feel and think about their lives.                                                                                                                                                                                                                                                                                                                                                                                                                                            | Indirectly Diener (1984) via MA              |
| D   | AWB and CWB are distinct constructs                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | Indirectly Lucas, Diener & Suh (1996) via MA |
| D1  | Affective well-being (AWB) refers to the presence of pleasant affect (e.g., feelings of happiness) and the absence of unpleasant affect (e.g., depressed mood).                                                                                                                                                                                                                                                                                                                                        | MA                                           |
| D2  | Cognitive well-being (CWB) refers to the cognitive evaluation of life overall (i.e., global life satisfaction) as well as of specific life domains (e.g., job satisfaction or marital satisfaction).                                                                                                                                                                                                                                                                                                   | MA                                           |
| E   | It is therefore plausible that other external life circumstances, such as life events, have differential effects on AWB and CWB and that adaptation of AWB and CWB does not occur at the same rate.                                                                                                                                                                                                                                                                                                    | MA                                           |
| F   | Life events have been examined from two major perspectives (Filipp & Aymanns, 2009): a stress perspective and a developmental perspective.                                                                                                                                                                                                                                                                                                                                                             | MA                                           |
| F1  | From the stress perspective, life events are viewed as specific types of stressors (e.g., Park, 2010; Segerstrom & Miller, 2004). Stressful life events are all events that significantly disturb the daily routine (Turner & Wheaton, 1997). This definition explicitly includes desirable events such as marriage or vacation (Holmes & Rahe, 1967).                                                                                                                                                 | MA                                           |
| F2  | By contrast, minor stressors such as daily hassles and uplifts (Kanner, Coyne, Schaefer, & Lazarus, 1981) are distinct from stressful life events. From the developmental perspective, life events are viewed as specific transitions. Transitions are defined as a “discontinuity in a person’s life space of which he is aware and which requires new behavioural responses” (Hopson & Adams, 1976, p. 24).                                                                                          | MA                                           |
| F3  | For the purpose of this metaanalysis, we developed a narrower definition of life events that integrates elements from both perspectives. According to our working definition, life events are time-discrete transitions that mark the beginning or the end of a specific status.                                                                                                                                                                                                                       | MA                                           |
| F4  | A status is a nominal variable with at least two levels.                                                                                                                                                                                                                                                                                                                                                                                                                                               |                                              |
| G   | the process perspective explicitly predicts the trajectory of SWB over time within individuals                                                                                                                                                                                                                                                                                                                                                                                                         | MA                                           |
| G1  | The adaptation process is initiated by an external stimulus (e.g., a major life event) that causes a physiological or psychological response (e.g., decreased SWB). Over time, the responsiveness diminishes and the level of SWB returns to its preevent level.                                                                                                                                                                                                                                       | MA                                           |
| H   | Most major life events are—at least to some extent— controllable and predictable and can therefore be anticipated. This anticipation might cause a specific hedonic reaction even before the event occurred: If the spouse is terminally ill, the hedonic reaction to bereavement starts long before the spouse actually dies. These so-called anticipatory or lead effects can be observed months or even several years before the occurrence of the event (Clark, Diener, Georgellis, & Lucas, 2008) | MA                                           |
| I   | The original treadmill theory suggested that people return to a neutral set point after an emotionally significant event. However, decades of research show that this part of the hedonic treadmill theory is wrong. Instead, most people are happy most of the time (Diener & Diener, 1996).                                                                                                                                                                                                          | TP                                           |
| J   | The empirical research that has been conducted since the publication of Brickman and Campbell (1971) reveals that if people do have set points, they vary considerably across individuals.                                                                                                                                                                                                                                                                                                             | TP                                           |
| K   | But we have found individual differences in the rate and extent of adaptation that occurs even to the same event. In our longitudinal studies, the size and even the direction of the change in life satisfaction varied considerably across individuals.                                                                                                                                                                                                                                              | TP                                           |
| L   | (not a direct citation) Set point of SWB = the level of SWB people return to after life events.                                                                                                                                                                                                                                                                                                                                                                                                        | TP                                           |
| M   | The difference between the current SWB and the set point affects the change rate. If the difference is larger, expected change is also larger.                                                                                                                                                                                                                                                                                                                                                         | OI(MA+TP)                                    |
| N   | (not a direct citation) People react differently to life events                                                                                                                                                                                                                                                                                                                                                                                                                                        | MA                                           |