### 2024-03-19 material/variable_definitions.Rmd

| **Construct** | **Scale Level** | **Range** | **Anchors** |
| --------- | ----------- | ----- | ------- |
| SWB | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
| SWBSetPoint | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
| LifeEvent | continous | [-1; 1] | -1 = objectively very negative \newline 0 = objectively neutral \newline 1 = objectively very positive |
| EmotionalReactivity | continuous | [0; 1] | 0 = emotional state is not influenced by external events \newline 1 = external events influence emotional states very strongly |
| InitialReaction | continuous | [-1; 1] | -1 = negative \newline 0 = neutral \newline 1 = positive |
| AdaptationRate | continuous | [0; 1] | 0 = no adaptation \newline 1 = fastest possible adaptation |
| AdaptationRatePerson | continous | [0; 1] | 0 = lowest possible tendency for adaptation \newline 1 = highest possible tendency for adaptation |